### 2020-08-27 Manuscript/What.lies.ahead.reviewed.tex

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions in both areas sampled, and a strong reduction in richness of current diatom assemblages. Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

@@

{\color{red}A growing body of evidence backs up the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine}}

@@

We characterized sample carbonate chemistry by measuring Total Alkalinity (TA; open-cell automated titration based on a 876 Dosimat plus (Metrohm AG) as part of a custom system assembled by Andrew Dickson (UCSD) and used in the laboratory of Alex Gagnon at UW) and Dissolved Inorganic Carbon (DIC; Apollo Instruments, USA; CO$_2$ extraction system with 10\% (v/v) phosphoric acid). Both measurements were calibrated and validated with certified reference material from the Scripps Oceanographic Institute. Using DIC and TA, we calculated pH and the remaining carbonate system parameters using the R package `seacarb` \cite{seacarb}.

@@

Code for all quality-screening and bioinformatics is available in the Supporting Information. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}. We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

@@

\subsection*{Year 2095 Environmental Scenario and Biological Responses}

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole. We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer''). To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017. See Supporting Information for details of sampling parameters.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, area, and season as predictor variables. We calculated the probability of each community, given these predictors, using the package `nnet` \cite{nnet2002}. Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year.

Within a geographic area and season, temperature and pH were most strongly associated with community-level changes. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects) {\color{red} -salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. We allowed} slope to vary by geographic area and intercept to vary by season, to capture large-scale seasonal differences associated with sunlight, photosynthesis, and related factors. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. See Supporting Information for analytical code.

Given the estimated sea-surface temperatures and pH values for 2095 and observed data from 2017, we then used taxon-specific logistic regression models to estimate the suitability of habitat for each taxon in the future scenario. For each point in the pH - temperature grid, we calculated species richness as the mean result of 100 independent draws of the posterior distributions of all modeled taxa, grouping the individual taxa models by phylum.

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (Red) and San Juan Island (Blue). \textbf{B} Planktonic richness (unique taxa) per sampling locality across seasons, as reflected by the eDNA COI assay. \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as facets.}}
\label{fig:Fig1}
\end{figure}
%%%%%

We collected intertidal environmental and genetic data from bottle samples each month between March 2017 and August 2018 (Figure 1A). Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with different points along the environmental gradient simultaneously showing differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

Metabarcoding analysis of eDNA samples generated more than 52.5M sequences from 745 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control (See Methods) the dataset included $\sim${}41M sequences, from 4782 unique amplicon sequence variants (ASVs). Of these, 1374 ASVs (22.8M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 222). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to season (Summer vs. Winter; PERMANOVA F-model = 14.4; p $<$ 0.001) and geographic Area (Hood Canal vs. San Juan Island; F-model = 15.4; p $<$ 0.001). We therefore performed a constrained analysis of principal components (CAP) for samples within each combination of Season and Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2 shows results for Hood Canal in summer as an example; full analysis in the Supporting Information). Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3 or 4) to identify groups of samples that were similar to one another with respect to biological community. Each biological cluster (colored hulls, Fig. 2) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2019-11-06.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

@@

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Figure3.20191105.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea. \textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions. Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.  \textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right). Median probabilities plotted. \textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}

>>>>

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions, and a strong reduction in richness of current diatom assemblages. Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

@@

{\color{red}A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}}

@@

We characterized sample carbonate chemistry by measuring Total Alkalinity (TA; open-cell automated titration based on a 876 Dosimat plus (Metrohm AG) as part of a custom system assembled by Andrew Dickson (UCSD) and used in the laboratory of Alex Gagnon at UW) and Dissolved Inorganic Carbon (DIC; Apollo Instruments, USA; CO$_2$ extraction system with 10\% (v/v) phosphoric acid). Both measurements were calibrated and validated with certified reference material from the Scripps Oceanographic Institute. Using DIC and TA, we calculated pH and the remaining carbonate system parameters using the R package `seacarb' \cite{seacarb}.

@@

Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database (CITE CRUX). We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

@@

\subsection*{Year-2095 Environmental Scenario and Biological Responses}

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole.   {\color{red}[ADD DETAIL HERE]}

%We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer''). To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017. See Supporting Information for details of sampling parameters.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables. {\color{red} Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models.} We calculated the probability of each community, given these predictors, using the package `nnet` \cite{nnet2002}. Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year.

Within a geographic area, temperature and pH were most strongly associated with community-level changes, although these conditions varied substantially between the Hood Canal and San Juan Island sites. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. Model selection using WAIC supported this as the preferred model over several similar ones (see Supporting Information).

Given the estimated sea-surface temperatures and pH values for 2095 and observed data from 2017, we then used taxon-specific logistic regression models to estimate the suitability of habitat for each taxon in the future scenario. For each point in the pH - temperature grid, we calculated species richness as the mean result of 100 independent draws of the posterior distributions of all modeled taxa.

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (Red) and San Juan Island (Blue). \textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay. \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as facets. {\color{red} CHANGE y-scale for Dinophyceae}}}
\label{fig:Fig1}
\end{figure}
%%%%%

We collected intertidal environmental and genetic data from bottle samples each month between March 2017 and August 2018 (Figure 1A). Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions. Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

Metabarcoding analysis of eDNA samples generated more than 52.5M sequences from 745 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control (See Methods) the dataset included $\sim${}41M sequences, from 4782 unique amplicon sequence variants (ASVs). Of these, 1374 ASVs (22.8M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 222). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B.1) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs. San Juan Island; F-model = 15.4; p $<$ 0.001). {\color{red} RPK: still up-to-date? no adonis code in the Rmd file that I saw} . We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2 shows results for Hood Canal in summer as an example; full analysis in the Supporting Information). Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3 or 4) to identify groups of samples that were similar to one another with respect to biological community. Each biological cluster (colored hulls, Fig. 2) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

{\color{red} RPK: I think everything below needs updating, according to some final analysis}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2020-08-12.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}
%Move this to the methods%
 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

@@

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-12.Other.sel.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea. \textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions. Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.  \textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right). Median probabilities plotted. \textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}
